---
title: "Reproducible document"
author: "Julia Prats"
format: html
editor: visual
---

## Basics of R

```{r}
1 + 1
```
---

## About me
- Júlia Prats, Center for Basic Metabolic Research, University of Copenhagen
- **Position**: PhD student
- *Proteomics*, *Skeletal Muscle*, *Metabolism*

---

## Simple code
```{r}
3 * 3
```
---

## Testing for Git
- I don't like chocolate 

---

## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

```
#with the message and warning we are telling R to avoid including all the junk information we don't need. Don't include the code junk but still run it. There are other common options are: include, echo, results, eval - look in the course information. 
#spaces are important in Rstudio
#csv reads data separated by ","
#csv2 reads data separated by ";"

## Showing the data 

```{r}
post_meal_data
```

# A nicer way of looking at the data is (according to Luke): 

```{r}
glimpse(post_meal_data)

```

#Commit; we would not commit the data (so the .csv file) but in this case it is very small so we will commit everything. Normally we would not commit the html either because it is a temporary file. 

#By looking at the data there are a few things that can be improved:
1.Too similar names
2.Too many IDs - what is the unique id
3.Decimal places
4.Order Ids
5.Different variables in one cell? Where are the units
6.There is two Fat.mass and two OGTT 
7.Capital letters that don't make sense
8.Separate datasets to avoid so many variables in one set


?colnames
help(colnames)

```{r}
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()

```



```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()
```
#You use "" when we are writing in natural language and no code language"@
#If you want to reference in the text just write @ in front of label

@fig-bmi-histo

## 9.5 Exercise: discrete plots

```{r}
#| fig-cap: "Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar

```{r}
#| fig-cap: "glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-bar

